---
title: "Bayesian Regression - Water Quality Samples"
output: md_document
---

**Author: Jacob Anderson**

```{r Load necessary packages}

### Load packages
### Check ArcGIS Pro license

library(arcgisbinding)
library(brms)
library(dplyr)
library(ggplot2)
library(ggspatial)
library(sf)
arc.check_product()
```

```{r Read the EPA Water Quality Indicators File and Set Detection Limit}

### Check the current directory

getwd()

### Read the csv and store as data frame

df_rec <- read.csv(
  "C:/Users/ja090/Desktop/Files/JHU/SU25_STATS/Water_Quality/water_quality_analysis/Data/RFEOutput2_v22_RRecs_20240301145200.csv")

### Inspect the structure of the data frame

head(df_rec)

### Detection Limit for Enterococcus in water samples
### CFU per 100 mL

DL <- 10
```

```{r Subset the Data}

### Subset the Data for the Enterococcus pollutant, at least 1 rejected samples, and for June 2023

df_subset <- subset(df_rec, Pollutant == "Enterococcus" & Rejected > 0 & YYYYMM == 202306)
```

```{r Check the size of your subset data frame}

### Use the dim function to retrieve the dimensions of the data frame subset

dim(df_subset)

### Dimensions: 1627 by 15
```

```{r Read the Station Location ID file}

### Read the csv and store as data frame

df_locs <- read.csv(
  "C:/Users/ja090/Desktop/Files/JHU/SU25_STATS/Quantitative_Assignment06/Files/RFEOutput2_v24_MonLocIdInfo_20240301145200.csv")

### Inspect the structure of the data frame

head(df_locs)
```

```{r Inner Join Two Dataframes (20 points)}

### Inner join the two tables with matching field (Monitoring Location ID)

df_subset_w_locs <- inner_join(
  df_subset,
  df_locs,
  by = "MonitoringLocationIdentifier"
)

### Inspect the structure of the data frame

head(df_subset_w_locs)
```

```{r Map the WQI Data}

### Convert the data frame into an sf data frame

sf_df <-st_as_sf(
  df_subset_w_locs,
  coords = c("LongitudeMeasure", "LatitudeMeasure"),
  crs = 4326
)

### Convert data frame to web mercator

sf_df_3857 <- st_transform(sf_df, 3857)

### Make a GGPLOT map of the sf data frame

ggplot() +
  annotation_map_tile(type = "osm") +
  geom_sf(
    data = sf_df_3857,
    color = "deeppink",
    size = 1
  ) +
  coord_sf(crs = 3857, expand = T) +
  labs(
    title = "Enterococcus Samples (Rejected)",
    x = NULL,
    y = NULL
  ) +
  theme_minimal()

```


```{r Define Data for the Bayesian Censoring}

### Observed data: detected values remain; censored values set to DL

df_subset["Censored Conc."] <- ifelse(df_subset$Maximum<DL, DL, df_subset$Maximum)

df_subset["Flag"] <- ifelse(df_subset$Maximum<DL, "left", "none")

df_subset["DetectionLimit"] <- DL
```

```{r Perform Data Censoring Using Stan}

### brms formula: | cens(Censored, DetectionLimit) ~ 1 means intercept-only model

fit <- brm(
  bf(Maximum | cens(Flag, DetectionLimit) ~ 1),
  data = df_subset,
  family = gaussian(),
  chains = 2, 
  iter = 2000,
  seed = 10
)

summary(fit)
```


```{r Visualizing the posterior}

### Convert brms fit to data frame

fit_df <- as_draws_df(fit)

### Plot the posterior probability distribution

ggplot(fit_df, aes(x = b_Intercept)) +
  geom_density(fill = "deeppink", alpha = 0.5) +
  labs(
    title = "Posterior Distribution of the Intercept",
    x = "Mean Enterococcus Concentration",
    y = "Density")

```

Using Bayesian censored regression with the brms package, we estimated the mean distribution of Enterococcus samples while accounting for left-censored values below the detection limit. The posterior distribution provides a full representation of uncertainty around this estimate, which would not be captured by a standard regression that ignores censoring.